A Generalized Reduced Linear Program for Markov Decision Processes

### 2016-07-08 single/errana.tex
Chandru's shortened version

%!TEX root =  autocontgrlp.tex
\section{Error Analysis}\label{sec:errana}
We now define the two projection operators which are central to our error analysis. The least upper bound (LUB) projection operator $\Gamma \colon \R^n \ra\R^n$ is defined below, see \eqref{gamdef}.
\begin{definition}\label{lubpop}
Given $J\in \R^n$ and the nonnegative valued vector $c\in \R^n_+$, define $r_{c,J}$ to be the solution to 
\label{lubplp}
 \underset{r\in \N'}{\min} &\,\, c^\top \Phi r\,,\\
 \text{s.t.}& \mb \Phi r\geq  TJ.
Then, for $J\in \R^n$, $\Gamma J$,
the least upper bound projection of $J$ is defined as 
\begin{align}\label{gamdef}
(\Gamma J)(i)\eqdef(\Phi r_{e_i,J})(i),\quad i=1,\ldots,n\,.
\end{definition}
\begin{remark}\label{ubrem}
To understand the meaning of $\Gamma$ (and $\hg$) define 
\begin{align}\label{ubclass}
\F_J\eqdef\{\,\Phi r\,:\,\Phi r\geq TJ, r\in \N\,\},
where $J\in \R^n$. Thus $\F_J$ contains vectors in the span of $\Phi$ that upper bound $TJ$. Further, since $(\Gamma J)(i) = \min\{ V(i) \,:\, V\in \F_J \}$, it also follows that $ V\ge \Gamma J $ holds for any $V\in \F_J$.
\end{remark}
The approximate least upper bound (ALUB) projection operator $\hg \colon \R^n \ra \R^n$ is defined as below:
\begin{definition}\label{alubpop}
Given $J\in \R^n$ and the nonnegative valued vector $c\in \R^n_+$, define $r'_{c,J}$ to be the solution to 
\begin{align}\label{alubplp}
 \underset{r\in \N'}{\min}& \,\mb c^\top \Phi r\,,\nn\\
 \text{s.t.}& \,\,\, W^\top E \Phi r\geq W^\top HJ.
Further, define $\hg J$  as 
\begin{align}\label{tgamdef}
(\hg J)(i)\eqdef(\Phi r'_{e_i,J})(i), \mb i=1,\ldots,n\,.
\end{definition}
\begin{lemma}\label{lpsol}
Let $A\in \R^{u\times v}$, $b,c\in \R^u$ and $b_0=Ax_0$ for 
some $x_0 \in \R^v$, $\N' \subset \R^v$ such that $\N' =x_0+ \N'$. Then
&\min\{c^\top Ax:Ax\geq b+b_0, x\in \N'\} \nn\\
&=\min\{c^\top A y:Ay \geq b, y \in \N' \}+c^\top b_0.
The claim follows by the change of variables $y := x-x_0$.
\noindent 
%Note that under our assumptions on $\N$, $\hg$ is well-defined.
%\begin{align*}
%r^*\eqdef\argmin_{r\in R^k}||J^*-\Phi r||_{\mn}\,.
%\end{align*}
%where $\psi$ is a Lyapunov function as in \cref{lyap}.
\begin{lemma}\label{bestbndmn}
||J^*-\Gamma J^*||_{\mn}\leq 2||J^*-\Phi r^*||_{\mn}.
Define $\eps = \norm{J^* - \Phi r^*}_{\mn}$ so that
$ J^*-\Phi r^* \le \eps \psi$ and $\Phi r^* - J^*\le \eps \psi$. Then from \cref{grlpassmp}-\eqref{ass:n4} it follows that $\Phi( r^* + \eps r_0 ) = \Phi r^* + \eps \psi \ge J^* = T J^*$. From the definition of $\Gamma$ in \eqref{gamdef} and \cref{ubrem}, we know that $\Phi r^*+\eps \psi \geq \Gamma J^*\geq J^*$. The result follows by noting that $2\eps\psi\geq \Phi r^*+\eps\psi -J^*\geq \Gamma J^*-J^*\geq 0$.
\begin{lemma}\label{tgmonotone}
For $J_1, J_2\in \R^n$ such that $J_1\leq J_2$, we have $\hg J_1\leq \hg J_2$.
Given $J\in \R^n$, let $\F_J\eqdef\{\,\Phi r\,: W^\top E \Phi r\geq W^\top HJ, r\in \N'\,\}\,$. Choose any $i\in \{1, \ldots, n\}$. Since $J_1\leq J_2$, from \cref{monotone} and \cref{grlpassmp}-\eqref{wassump} it follows that $W^\top H J_1\leq W^\top H J_2$. Hence, $\F_{J_2} \subset \F_{J_1}$ and thus $(\hg J_1)(i) \le (\hg J_2)(i)$.  Since $i$ was arbitrary, the result follows. 

@@

\hg( J + t \psi) \le \hg J + \beta t \psi,

@@

$T$ is a $\norm{\cdot}_{\mn}$-contraction with factor $\beta_{\psi}$.

@@

By definition, for $1\le i \le n$, $(\hg (J+t\psi) )(i) = \min\{ e_i^\top \Phi r \,:\, W^\top E\Phi r \ge W^\top H(J+t\psi), r\in \N' \}$.
By \eqref{eq:psilin}, as $t>0$, $H(J+t\psi) \le HJ + t \beta_\psi \psi$ and hence $W^\top H(J+t\psi) \le W^\top (HJ + t \beta_\psi \psi)$. Thus,

@@

 \min\{ e_i^\top \Phi r \,:\, W^\top E\Phi r \ge W^\top HJ+t\beta_\psi \psi), r\in \N' \}$.
Now, using Lemma~\ref{lpsol} with $A=W^\top E \Phi$, $b=W^\top HJ$, $c=e_i$, $b_0=t\beta_\psi \psi$ 
and $x_0=t \beta_\psi r_0$, the statement follows since $A x_0 = b_0$ (from \cref{grlpassmp}-\eqref{lyap}) and $\N' = \N' + \alpha t r_0$ (from \cref{grlpassmp}-\eqref{ass:n4}).

@@

In what follows we denote by $\hv$ the unique fixed point of $\hg$ i.e., $\hv = \hg \hv\,$. 
We now show that the vector $\hj$ dominates the vector $\hv$:
\begin{lemma}\label{relation1}

@@

It suffices to show that $V_1\eqdef \hg V_0 \le V_0 \le \hj$ since then the desired result follows
by defining $V_{n+1} = \hg V_n$, $n\ge 1$, noting that by \cref{tgmonotone}, $V_{n+1}\le V_{n}$ and by  \cref{tmaxnormmn}, $V_n \to \hv$.

@@

We need to show that $V_1(i)=(\hg V_0)(i) = (\Phi r'_{e_i,V_0})(i) \leq V_0(i)$. 
By the definition of $r'_{e_i,V_0}$ we know that $(\Phi r'_{e_i,V_0})(i) \le (\Phi r)(i)$

@@

Hence, by the monotone property of $H$ and \cref{grlpassmp}-\eqref{wassump} it is sufficient if $\Phi r_i \ge V_0$.

@@

The vectors $\hv,\tj$ obey $\tj\geq\hv$.
Let $ r_1,  r_2,\ldots, r_n$ be solutions to the ALP in \eqref{alp} (with an additional constraint that the solution be restricted to $\N$) for $c=e_1, e_2,\ldots,e_n$, respectively,
and define $V_0\in \R^n$ by $V_0(i)=\underset{j=1,\ldots,n}{\min}(\Phi r_j)(i)$, $1\le i \le n$. The rest of the proof 
follows in the same manner as the proof of \cref{relation1}.

@@

\min_{r\in \N}\, &||\Phi r-\hv||_{1,c}\\
\text{s.t.}\mb &\, W^\top E \Phi r\geq W^\top H \Phi r.

@@

\begin{lemma}\label{srwalp}
$\tr \in \R^k$ is a solution to ALP \eqref{alp} iff it solves the following program:
\text{s.t.}\mb &\, \Phi r\geq  T \Phi r.

@@

We know from Lemma~\ref{relation2} that $\tj = \Phi \tr \geq\hv$, and the rest of the arguments follow in the same manner as proof for \cref{srw}.
||J^*-\hv||_{\mn}&\leq \frac{1}{{1-\beta_{\psi}}}\big(2||J^*-\Phi r^*||_{\mn}&\nn\\+||\Gamma J^*-\hg J^*||_{\mn}\big).
Recall that $\hg \hv=\hv$. By the triangle inequality,

@@

&||J^*-\hv||_\mn\\&\leq ||J^*-\hg J^*||_\mn+||\hg J^*-\hg \hv||_\mn\\&\leq ||J^*-\hg J^*||_\mn+\beta_\psi || J^*- \hv||_\mn,

@@

and so by reordering and with another triangle inequality,

@@

&\norm{J^*-\hv}_{\mn} \nn\\&\le \frac{\norm{ J^*-\hg J^*}_{\mn}}{1-\beta_\psi}\nn\\

@@

The proof follows by applying \cref{bestbndmn} on \eqref{jhv}.

@@

For this result, recall that $r_0 \in \R^k$ is the vector such that $\psi = \Phi r_0$ by assumption.

@@

\begin{align}\label{part}
&||\Phi r'-J^*||_{\mn}\nn\\&\leq ||\Phi r^*-J^*||_{\mn}+||\Phi r'-\Phi r^*||_{\mn}\nn\\

@@

&=\frac{2}{1-\beta_\psi}\gamma.
%We know that $\tr \in \N$ by \cref{grlpassmp}-\eqref{nassmp} and 
$r'$ is feasible for the ALP in \eqref{grlp} by \cref{restate}. Then from \cref{srwalp} it follows that
%As a result, from \cref{srw} it follows that

@@

&||\hj-\hv||_{1,c}\leq||\Phi \tr-\hv||_{1,c}\leq||\Phi r'-\hv||_{1,c}\nn\\

@@

The result follows from \cref{cmt1mn} and \eqref{part}.

@@

We now bound the performance of the greedy policy $\hu$.
\begin{theorem}[Control Error Bound in $\norm{\cdot}_{\mn}$]
\label{polthe}
Let $\hu$ be the greedy policy with respect to the solution $\hj$ of the GRLP and $J_{\hu}$ be its value function.
% Let $r^*$ be as in Theorem~\ref{mt2mn}, then
Then,
\begin{align}\label{polthebnd}
%||J_{\hu}-\hj||_{1,c}
\norm{J^* - J_{\hu}}_{1,c}
&\leq 2\left(\frac{c^\top \psi}{1-\beta_{\psi}}\right)^2\, \big(6 ||J^*-\Phi r^*||_{\mn}
\nn\\&
+2||\Gamma J^*-\hg J^*||_{\mn}\big).
||J^*-J_{\hu}||_{1,c}&\leq ||J^*-\hj||_{1,c}+||J_{\hu}-\hj||_{1,c}\,.
Let us now bound the second term on the right-hand side.
Since $\hu$ is greedy w.r.t. $\hj$, it holds that $T_{\hu} \hj = T \hj$.
Also, $T_{\hu} J_{\hu} = J_{\hu}$.
Hence, $J_{\hu} - \hj = T_{\hu} J_{\hu} - T_{\hu} \hj + T \hj - \hj
=\alpha P_{\hu} (J_{\hu}- \hj) + T\hj - \hj$.
Hence,
\begin{align}\label{polderv}
&||J_{\hu}-\hj||_{1,c}\nn\\&=||(I-\alpha P_{\hu})^{-1}(T\hj-\hj)||_{1,c}\nn\\
&\leq c^\top(I-\alpha P_{\hu})^{-1}|T\hj-\hj|\nn\\
&\leq c^\top (I-\alpha P_{\hu})^{-1} \psi ||T\hj-\hj||_{\mn}\nn\\
&\leq \frac{c^\top \psi}{1-\beta_{\psi}}||T\hj-\hj||_{\mn}\nn\\
%&\leq \frac{c^\top \psi}{1-\beta_{\psi}}||T\hj-TJ^* +J^*- \hj||_{\mn}\nn\\
&\leq \frac{c^\top \psi}{1-\beta_{\psi}}(||T\hj-TJ^*||_{\mn} +||J^*- \hj||_{\mn})\nn\\
&\leq \frac{c^\top \psi}{1-\beta_{\psi}}(1+\beta_{\psi})||J^*- \hj||_{\mn},
where in the second inequality, we use Jensen's inequality and $|T\hj - \hj|$ stands for the 
vector whose $i$th component is $|(T\hj)(i) - \hj(i)|$. Further, the last inequality follows
since $T$ is a $\norm{\cdot}_{\mn}$ contraction with factor $\beta_{\psi}$ as noted earlier.
%componentwise  $(I-\alpha P_{\hu})^{-1}$ is a positive operator for $x=(x_1,\ldots,x_n)^\top\in \R^n$, $|x|=(|x_1|,\ldots,|x_n|)^\top\in \R^n$. 
Hence,
&||J^*-J_{\hu}||_{1,c}\nn\\
%&\leq ||J^*-\hj||_{1,c}+||J_{\hu}-\hj||_{1,c}\nn\\
&\leq c^\top \psi ||J^*-\hj||_{\mn}+c^\top \psi\frac{1+\beta_\psi}{1-\beta_\psi}||J^*- \hj||_{\mn}\nn\\
&=\frac{2c^\top \psi}{1-\beta_{\psi}}||J^*- \hj||_{\mn}.
The result now follows by substituting the bound on $||J^*- \hj||_{\mn}$ from \cref{cmt2mn}.
\begin{note}
By bounding $\etmn=||\Gamma J^*-J^*+J^*-\hg J^*||_{\mn}\leq 2||J^*-\Phi r^*||_{\mn}+||J^*-\hg J^*||_{\mn}$ 
(the inequality follows from \cref{bestbndmn}), 
we can loosen the bounds in \cref{cmt2mn} and \cref{polthe} to
\label{loose1}
||J^*-\hj||_{1,c}&\leq\frac{c^\top\psi}{1-\beta_\psi}(10 ||J^*-\Phi r^*||_{\mn}
\nn\\&
+2||J^*-\hg J^*||_{\mn}).\\
\label{loose2}
||J^* - J_{\hu}||_{1,c}&\leq 2\left(\frac{c^\top \psi}{1-\beta_{\psi}}\right)^2 \,\big(10 ||J^*-\Phi r^*||_{\mn}
\nn\\&
+2||J^*-\hg J^*||_{\mn}\big).
Here the term $||J^*-\hg J^*||$ in \eqref{loose1} and \eqref{loose2} captures the error due to the use of both $\Phi$ and $W$. Though, \eqref{loose1} and \eqref{loose2} might be loser bounds than \eqref{finalbndmn} and \eqref{polthebnd} respectively, the advantage of this bound is that it captures the error due to function approximation as well as constraint reduction in a direct manner.
\end{note}

>>>>

%!TEX root =  autocontgrlp.tex
\section{Error Analysis}\label{sec:errana}
The least upper bound (LUB) projection operator $\Gamma \colon \R^n \ra\R^n$ is defined as below:
\begin{definition}\label{lubpop}
Given $J\in \R^n$ and the nonnegative valued vector $c\in \R^n_+$, define $r_{c,J}$ to be the solution to 
\begin{align}
\label{lubplp}
\begin{split}
 \underset{r\in \N'}{\min} &\,\, c^\top \Phi r\,,\\
 \text{s.t.}& \mb \Phi r\geq  TJ.
 \end{split}
\end{align}
Then, for $J\in \R^n$, $\Gamma J$,
the least upper bound projection of $J$ is defined as 
\begin{align}\label{gamdef}
(\Gamma J)(i)\eqdef(\Phi r_{e_i,J})(i),\quad i=1,\ldots,n\,.
\end{align}
\end{definition}
The approximate least upper bound (ALUB) projection operator $\hg \colon \R^n \ra \R^n$ is defined as below:
\begin{definition}\label{alubpop}
Given $J\in \R^n$ and the nonnegative valued vector $c\in \R^n_+$, define $r'_{c,J}$ to be the solution to 
\begin{align}\label{alubplp}
 \underset{r\in \N'}{\min}& \,\mb c^\top \Phi r\,,\nn\\
 \text{s.t.}& \,\,\, W^\top E \Phi r\geq W^\top HJ.
\end{align}
Then, for $J\in \R^n$ define $\hg J$  as 
\begin{align}\label{tgamdef}
(\hg J)(i)\eqdef(\Phi r'_{e_i,J})(i), \mb i=1,\ldots,n\,.
\end{align}
\end{definition}
\begin{assumption}\label{ass:n4}
The set $\N'$ is such that $\N' = \N + t r_0$ for any $t\in \R$, where $r_0\in \R^k$ such that $\Phi r_0 = \psi$ and $\N$ is as in \eqref{grlp} satisfying \cref{nassmp}.
\end{assumption}
\begin{lemma}\label{lpsol}
Let $A\in \R^{u\times v}$, $b,c\in \R^u$ and $b_0=Ax_0$ for 
some $x_0 \in \R^v$, $\N' \subset \R^v$ such that $\N' =x_0+ \N'$. Then
\begin{align}
&\min\{c^\top Ax:Ax\geq b+b_0, x\in \N'\} \nn\\
&=\min\{c^\top A y:Ay \geq b, y \in \N' \}+c^\top b_0.
\end{align}
\end{lemma}
\begin{proof}
The claim follows by the change of variables $y := x-x_0$.
\end{proof}
\noindent 
%Note that under our assumptions on $\N$, $\hg$ is well-defined.
%\begin{align*}
%r^*\eqdef\argmin_{r\in R^k}||J^*-\Phi r||_{\mn}\,.
%\end{align*}
%where $\psi$ is a Lyapunov function as in \cref{lyap}.
\begin{lemma}\label{bestbndmn}
We have
\begin{align}
||J^*-\Gamma J^*||_{\mn}\leq 2||J^*-\Phi r^*||_{\mn}.
\end{align}
\end{lemma}
\begin{proof}
Define $\eps = \norm{J^* - \Phi r^*}_{\mn}$ so that
$ J^*-\Phi r^* \le \eps \psi$. Then from \cref{ass:n4} it follows that $\Phi( r^* + \eps r_0 ) = \Phi r^* + \eps \psi \ge J^* = T J^*$. From the definition of $\Gamma$ in \eqref{gamdef} we know that $\Phi r^*+\eps \psi \geq \Gamma J^*\geq J^*$. The result follows by noting that $2\eps\psi\geq \Phi r^*+\eps\psi -J^*\geq \Gamma J^*-J^*\geq 0$.
\end{proof}
\begin{lemma}\label{tgmonotone}
For $J_1, J_2\in \R^n$ such that $J_1\leq J_2$, we have $\hg J_1\leq \hg J_2$.
\end{lemma}
\begin{proof}
Given $J\in \R^n$, let $\F_J\eqdef\{\,\Phi r\,: W^\top E \Phi r\geq W^\top HJ, r\in \N\,\}\,$. Choose any $i\in \{1,\ldots,n\}$. Since $J_1\leq J_2$, from \cref{wassump} it follows that $W^\top H J_1\leq W^\top H J_2$. Hence, $\F_{J_2} \subset \F_{J_1}$ and thus $(\hg J_1)(i) \le (\hg J_2)(i)$.  Since $i$ was arbitrary, the result follows. 
\end{proof}

%Since most of our analysis in sections~\ref{sec:lubp} and ~\ref{sec:alubp} depended on showing that $\Gamma$ is a contraction map in the $L_\infty$ norm, we first show that $\Gamma$ is also a contraction map in the weighted $L_\infty$ norm.
%The generalization of  \cref{maxnorm} for the weighted $L_{\infty}$ norm is as follows:

@@

\hg( J + t \psi) \le \hg J + \beta t \psi\,.

@@

Let us note in passing that from this result and~\eqref{eq:psilin} it immediately follows that $T$ is an $\norm{\cdot}_{\mn}$-contraction with factor $\beta_{\psi}$.

@@

By definition, for $1\le i \le n$, $(\hg (J+t\psi) )(i) = \min\{ e_i^\top \Phi r \,:\, W^\top E\Phi r \ge W^\top H(J+t\psi), r\in \N \}$.
By \cref{eq:psilin}, as $t>0$, $H(J+t\psi) \le HJ + t \beta \psi$ and hence $W^\top H(J+t\psi) \le W^\top (HJ + t \beta \psi)$. Thus,

@@

 \min\{ e_i^\top \Phi r \,:\, W^\top E\Phi r \ge W^\top H(J+t\psi), r\in \N \}$.
Now, using Lemma~\ref{lpsol} with $A=W^\top E \Phi$, $b=TJ$, $c=e_i$, $b_0=\beta t \psi$ 
and $x_0=t \beta r_0$, the statement follows
as thanks to \cref{lyap}, $A x_0 = b_0$ and thanks to
\cref{ass:n4}, $\N = \N + \alpha t r_0$.

@@

In what follows we denote by $\hv$ the unique fixed point of $\hg$ i.e., $\hv = \hg \hv\,$. We now show that the vector $\hj$ dominates the vector $\hv$:

@@

It suffices to show $V_1\eqdef \hg V_0 \le V_0 \le \hj$ since then the desired result follows
by defining $V_{n+1} = \hg V_n$, $n\ge 1$, noting that by \cref{tgmonotone}, $V_{n+1}\le V_{n}$ and by  \cref{tmaxnormmn}, $V_n \to \hv$.

@@

We need to show that $V_1(i)=(\hg V_0)(i) = (\Phi \har_{e_i,V_0})(i) \leq V_0(i)$. 
By the definition of $\har_{e_i,V_0}$ we know that $(\Phi \har_{e_i,V_0})(i) \le (\Phi r)(i)$

@@

Hence, by the monotone property of $H$ and \cref{wassump} it is sufficient if $\Phi r_i \ge V_0$.

@@

%The following lemmas relate the fixed point $\hv$ of $\hg$ to the solution $\hj$ of the GRLP in \eqref{grlp}.

@@

\text{s.t.}\mb &\, W^\top E \Phi r\geq W^\top H \Phi r\,, \qquad r\in \N\,.

@@

||J^*-\hv||_{\mn}\leq \frac{2||J^*-\Phi r^*||_{\mn}+||\Gamma J^*-\hg J^*||_{\mn}}{1-\beta_{\psi}}.

@@

@@

&||J^*-\hv||_\mn\\&\leq ||J^*-\hg J^*||_\mn+||\hg J^*-\hg \hv||_\mn\\&\leq ||J^*-\hg J^*||_\mn+\beta_\psi || J^*- \hv||_\mn

@@

and so by reordering and another triangle inequality,

@@

&\norm{J^*-\hv}_{\mn} \\&\le \frac{\norm{ J^*-\hg J^*}_{\mn}}{1-\beta_\psi}\\

@@

The proof is complete by using \cref{bestbndmn} in \eqref{jhv}.

@@

For this result, recall that $r_\in \R^k$ is the vector such that $\psi = \Phi r_0$ by assumption.

@@

&||\Phi r'-J^*||_{\mn}\\&\leq ||\Phi r^*-J^*||_{\mn}+||\Phi r'-\Phi r^*||_{\mn}\nn\\

@@

&=\frac{2}{1-\beta_\psi}\gamma.\nn
Since $r'$ is also feasible for the GRLP in \eqref{grlp} we have

@@

&||\hj-\hv||_{1,c}\leq||\Phi r'-\hv||_{1,c}\nn\\

@@

The result follows from \cref{cmt1mn}.

@@
